Modeling the competing effects of the immune system and EMT on epithelial cancers

### 2019-12-10 Paper/CleanPaper.tex
update res/discussion

\linenumbers

@@

The roles that TGF-$\beta$ play throughout the tumor and its microenvironment also warrant further investigation. We found that, below a certain threshold, reduction of TGF-$\beta$ increases the Time to Cancer  (Fig. \ref{fig:FirstSurvivalCurves}E), thus reducing expression levels of TGF-$\beta$ in the tumor microenvironment benefits survival. Intriguingly, recent experimental work demonstrated that TGF-$\beta$ drives tumor suppression in pancreatic cancer by promoting EMT \cite{david16_tgfv}. However, TGF-$\beta$ is a master regulator implicated in numerous cellular signaling processes, and changing the concentrations of TGF-$\beta$ even in a local tumor microenvironment could have large off-target effects. Indeed, it has been shown that TGF-$\beta$ promotes invasion and heterogeneity (although suppresses cell proliferation) in squamous cell carcinoma \cite{oshimori15_tgfv}. Future work thus ought to consider the effects of targeting signaling factors downstream of TGF-$\beta$ that still have the ability to modulate epithelial cell dynamics. Towards this end, we are currently developing a larger TGF-$\beta$ signaling pathway module with appropriate crosstalks to  epithelial/mesenchymal/immune cell functions to be incorporated into the model.

@@

A further goal for future work is to explore (and exploit) the heterogeneity of tumor evolution in greater depth: this heterogeneity aids the evasion of the tumor from immune effects. Studying the consequences of decanalization \cite{gibson09_decanalization} during cancer progression is too-often sidelined, despite evidence supporting its prominence \cite{cyll17_tumour, punt17_tumour, dagogo-jack18_tumour}. Yet despite these challenges, for which the complexity of the disease may be often in large part responsible, great progress has been and continues to be made. As we approach a new generation of immunotherapies, it is these very complexities that we must better understand in order to control or eradicate the disease.

>>>>

%\linenumbers

@@

The competing roles that TGF-$\beta$ play throughout the tumor and its microenvironment also warrant further investigation. We found that -- below a certain threshold -- reduction of TGF-$\beta$ increases the time to invasion (Fig. \ref{fig:FirstSurvivalCurves}E): reducing  TGF-$\beta$ in the TME can benefit survival. Recent experiment work in support of this demonstrated that TGF-$\beta$ drives tumor suppression in pancreatic cancer by promoting EMT \cite{david16_tgfv}. TGF-$\beta$, however, is implicated in numerous other cellular signaling processes, and changing TGF-$\beta$ concentration even in a local environment could have large off-target effects. Indeed, it has been shown that TGF-$\beta$ promotes invasion and heterogeneity while  suppressing cell proliferation in squamous cell carcinoma \cite{oshimori15_tgfv}. To help account for such complex circuitry, future work should incorporate the effects of signaling factors downstream of TGF-$\beta$ on tumor cell dynamics. We are currently developing a larger TGF-$\beta$ signaling pathway component within the full model that permits crosstalks between ETCs, MTCs, and immune cell populations through transcriptional signaling activity.

@@

Further {\em in silico} work on this model and its extensions will explore (and exploit) the heterogeneity of tumor evolution in greater depth. Tumor heterogeneity greatly enhances the capabilities of the tumor to evade immune effects. Studying the consequences of this increased heterogeneity following disease incidence, i.e. decanalization \cite{gibson09_decanalization}, is too-often sidelined, despite mounting evidence in support of its prominent role in cancer evolution \cite{cyll17_tumour, punt17_tumour, dagogo-jack18_tumour}. Despite these challenges (for which disease complexity is often in part responsible), great progress has been and continues to be made. As we are rapidly approaching a new generation of immunotherapies, it is these very complexities that we must better understand in order to control or eradicate the disease.
